---
title: "Uma Maheshwari C"
subtitle: Assignment 1
date: "2024-02-02"
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

```{r}
library(readr)
processes2 <- read_csv("C://Users//ujwal//Downloads//iris.csv")
View(processes2)
```
#Print out descriptive statistics for a selection of quantitative and categorical variables.

# Define quantitative variables
```{r}
variable_1 <- c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")
```

# Print out descriptive statistics for quantitative variables
```{r}
quant_desc <- summary(iris[, variable_1])
print("Descriptive statistics for quantitative variables:")
print(quant_desc)
```

# Define categorical variable
```{r}
variable_2 <- "Species"
```

# Print out descriptive statistics for categorical variable
```{r}
cat_counts <- table(iris[[variable_2]])
cat(paste("Descriptive statistics for", variable_2, ":"))
cat("\n")
print(cat_counts)
```

# Define the variable to transform
```{r}
vars_to_trans <- "Sepal.Width"
```

# Perform the transformation (squaring each value)
```{r}
iris_trans <- iris
iris_trans[[vars_to_trans]] <- iris[[vars_to_trans]]^2
head(iris_trans)
```

```{r}
library(ggplot2)
```
# Plot histogram of Petal.Length
```{r}
hist <- ggplot(iris, aes(x = Petal.Length)) +
  geom_histogram(binwidth = 0.4, fill = "blue", color = "black") +
  labs(title = "Histogram of Petal Length",
       x = "Petal Length",
       y = "Frequency")
print(hist)
```

# Create scatterplot of Petal.Length vs Petal.Width

```{r}
scatplot <- ggplot(iris, aes(x = Petal.Length, y = Petal.Width)) +
  geom_point(color = "red") +
  labs(title = "Scatterplot of Petal Length vs Petal Width",
       x = "Petal Length",
       y = "Petal Width")
print(scatplot)
```
