---
title: "FML Assignment 2"
author: "Uma Maheshwari C"
date: "2024-02-25"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
Purpose
The purpose of this assignment is to use k-NN for classification.

#Directions
Universal bank is a young bank growing rapidly in terms of overall customer acquisition.
The majority of these customers are liability customers (depositors) with varying sizes of
relationship with the bank. The customer base of asset customers (borrowers) is quite
small, and the bank is interested in expanding this base rapidly in more loan business. In
particular, it wants to explore ways of converting its liability customers to personal loan
customers.
A campaign that the bank ran last year for liability customers showed a healthy conversion
rate of over 9% success. This has encouraged the retail marketing department to devise
smarter campaigns with better target marketing. The goal is to use k-NN to predict whether
a new customer will accept a loan offer. This will serve as the basis for the design of a new
campaign.
The file UniversalBank.csv contains data on 5000 customers. The data include customer
demographic information (age, income, etc.), the customer’s relationship with the bank
(mortgage, securities account, etc.), and the customer response to the last personal loan
campaign (Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the
personal loan that was offered to them in the earlier campaign.

***
#Libraries
```{r}
library(class)
library(caret)
library(e1071)
```
# Given data set
```{r }
universal.df <- read.csv("C://Users//ujwal//Downloads//UniversalBank.csv")
dim(universal.df)
t(t(names(universal.df))) 
```
Drop ID and ZIP
```{r}
universal.df <- universal.df[,-c(1,5)]
```

Partition the data into training (60%) and validation (40%) sets. Before we split, let us transform categorical variables into dummy variables
```{r}
# considering education as a factor
universal.df$Education <- as.factor(universal.df$Education)

# creating dummy variables

groups <- dummyVars(~., data = universal.df) 
universal_m.df <- as.data.frame(predict(groups,universal.df))


set.seed(1)  # setting seed for reproducible
train.index <- sample(row.names(universal_m.df), 0.6*dim(universal_m.df)[1])
valid.index <- setdiff(row.names(universal_m.df), train.index)  
train.df <- universal_m.df[train.index,]
valid.df <- universal_m.df[valid.index,]
t(t(names(train.df)))


```
Normalizing the data is mandatory
```{r}
train.norm.df <- train.df[,-10] 
valid.norm.df <- valid.df[,-10]

norm.values <- preProcess(train.df[, -10], method=c("center", "scale"))
train.norm.df <- predict(norm.values, train.df[, -10])
valid.norm.df <- predict(norm.values, valid.df[, -10])
```

### Questions

Consider the following customer:

1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember to transform categorical predictors with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff value of 0.5. How would this customer be classified?

```{r}
# creating new sample
new_customer <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)

# Normalize the new customer data
new.cust.norm <- new_customer
new.cust.norm <- predict(norm.values, new.cust.norm)

```
Predict using knn classification method
```{r}

knn.pred1 <- class::knn(train = train.norm.df, 
                       test = new.cust.norm, 
                       cl = train.df$Personal.Loan, k = 1)
knn.pred1

```
2. What is a choice of k that balances between overfitting and ignoring the predictor information?
```{r}
# Calculate the accuracy for each value of k
# Set the range of k values to consider

accuracy.df <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
  knn.prediction <- class::knn(train = train.norm.df, 
                         test = valid.norm.df, 
                         cl = train.df$Personal.Loan, k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.prediction, 
                as.factor(valid.df$Personal.Loan),positive = "1")$overall[1]
  
}

which(accuracy.df[,2] == max(accuracy.df[,2])) 
```
3. Show the confusion matrix for the validation data that results from using the best k.
```{r}

knn.valid <- knn(train = train.df[,-10],
             test = valid.df[,-10],
             cl = train.df$Personal.Loan,
             k=3)
knn.valid
```

```{r}
confusionMatrix(knn.valid,as.factor(valid.df[, 10]))
```

4. Consider the following customer: Age = 40, Experience = 10, Income = 84,
Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,
Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
Card = 1. Classify the customer using the best k.

```{r}
customer.df= data.frame(Age = 40, 
                        Experience = 10, 
                        Income = 84,
                        Family = 2, 
                        CCAvg = 2, 
                        Education_1 = 0, 
                        Education_2 = 1, 
                        Education_3 = 0, 
                        Mortgage = 0, 
                        Securities.Account = 0, 
                        CD.Account = 0,
                        Online = 1, 
                        CreditCard = 1)
knn.pred2 <- knn(train = train.df[,-10],
             test = customer.df, 
             cl = train.df[,10], 
             k=3, prob=TRUE)
knn.pred2
```
5.Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on the differences and their reason.

```{r}
## Already we converted Education to Dummy Variables initially
set.seed(123) #setting seed to 123 as we are dividing it into training, validation and testing sets

train.index <- sample(rownames(universal_m.df), 0.5*dim(universal_m.df)[1]) # 50%

valid.index <- sample(setdiff(rownames(universal_m.df),train.index),
                      0.3*dim(universal_m.df)[1]) # 30%

test.index = setdiff(rownames(universal_m.df), union(train.index, valid.index))

train.df<- universal_m.df[train.index, ]
valid.df <- universal_m.df[valid.index, ]
test.df <- universal_m.df[test.index, ]

head(train.df)
head(valid.df)
head(test.df)

```
```{r}
dim(train.df)
dim(valid.df)
dim(test.df)
```


```{r}
#normalizing the above values

train.norm.df <- train.df[,-10] 
valid.norm.df <- valid.df[,-10]
test.norm.df <- test.df[,-10]


norm.values <- preProcess(train.df[, -10], method=c("center", "scale"))
train.norm.df <- predict(norm.values, train.df[, -10])
valid.norm.df <- predict(norm.values, valid.df[, -10])
test.norm.df <- predict(norm.values, test.df[, -10])

knn_prediction4 <- class::knn(train = train.norm.df, test = train.norm.df, cl = train.df$Personal.Loan, k= 3)
knn_prediction4
```

```{r}
confusion_matrix1 <- confusionMatrix(knn_prediction4, as.factor(train.df$Personal.Loan))
confusion_matrix1

```
```{r}
knn_pred <- class::knn(train = train.norm.df,
test = valid.norm.df,
cl= train.df$Personal.Loan, k= 3)
knn_pred

```
```{r}
confusion_matrix2 <- confusionMatrix(knn_pred, as.factor(valid.df$Personal.Loan))
confusion_matrix2
```
```{r}
knn_pred6 <- class::knn(train = train.norm.df,
test = test.norm.df,
cl= train.df$Personal.Loan, k= 3)
knn_pred6
```
```{r}
confusion_matrix3 <- confusionMatrix(knn_pred6, as.factor(test.df$Personal.Loan))
confusion_matrix3
```
Differentiating test confusion matrix of the training and validation sets.

1. Predictions are less accurate since the test model's accuracy is lower than that of the training and validation sets.
2. The test case's Kappa is low. 
3. The test case's sensitivity is likewise low in comparison to them, indicating a poor capacity to recognize affirmative cases.
4. Specificity is likewise low, indicating a poor capacity to recognize negative cases.
Overall, the test case confusion matrix is less than theirs; the variations may be caused by variables such as data set selection, parameter setup, and approach strategy. 
